Using Random Walk Simulations to Calculate Ground State Energies in Quantum Physics

### 2020-04-19 Paper/main.tex
Rewrote Method section based on feedback

\begin{equation}
  \label{eq:squarewell}
    a(j) =
    \begin{cases}
      0,& \text{if } j < J\\
      \infty,& \text{if } j \geq J
    \end{cases}
    \nonumber
\end{equation}
where J is a well-defined boundary. So the walker cannot die inside the well but
wil immediately die when they step outside, which we expect will definitely
happen given sufficient time. Inside the well, the potential is simply that of a
free particle:

@@

\section{Method}

The Random walk is implemented computationally \todo{cite} and is extensible
to multiple dimensions. For a $N$-dimensional lattice, at each point there are
$2N$ possible directions for the walker to move. So a direction is selected
randomly using a Mersenne Twister pseudorandom number generator (pRNG). This algorithm
was chosen as it is a very fast implementation of a pRNG \todo{cite} and it
passes many statistical randomness tests \todo{cite}, and has a very large
period of $2^{19937-1}$.

The random walk ends when a walker dies, with a
specifically chosen $a(j)$. Initially, the $a(j)$ is chosen to be that of an
infinite square well \todo{cite}, that is:
\begin{equation}
  \label{eq:squarewell}
    a(j) =
    \begin{cases}
      0,& \text{if } j < J\\
      \infty,& \text{if } j \geq J
    \end{cases}
    \nonumber
\end{equation}
where J is a well-defined boundary. So the walker has no chance of dying whilst
within the well, but will die immediately once reaching the boundary. As
explained in Section \ref{sec:TheoreticalBackground}, we expect that given
sufficient time, all the walkers will reach the boundary and die. But in order
to save computational resources, a maximum number of steps may also be defined,
such that a walker will also die should they move that number of steps without
reaching a boundary. This was not initially implemented, as it can introduce
systematic errors, since there would be a disproportionate number of walkers dying
on the maximum number of steps.

Thus, the Schr\"{o}dinger equation inside a one-dimensional infinite square well
is simply that of a free particle \todo{cite}:
\begin{equation}
  \frac{-\hbar^2}{2m}\frac{d^2 \psi}{dx^2} = E\psi(x)
  \nonumber
\end{equation}
where $x$ is the position of the walker. By making the substitution $x = J
\times j/R$, such that $j=R$ corresponds to the walker being at the boundary, we
obtain:
\begin{equation}
  \begin{split}
    -\frac{\hbar^2}{2m}\frac{R^2}{J^2}\frac{d^2\psi}{dj^2} &= E\psi(x)\\
    -\frac{1}{2}\frac{d^2\psi}{dj^2} &= \frac{mJ^2E}{\hbar^2R^2}\psi(x)
  \end{split}
  \nonumber
\end{equation}

If we compare this to Equation \ref{eq:randomwave}, allowing $a(j) = 0$ inside
the well, we see that:
\begin{equation}
  \begin{split}
    \lambda &= \frac{mJ^2E}{\hbar^2R^2} \\
    E &= \lambda R^2 \frac{\hbar^2}{mJ^2}
  \end{split}
  \nonumber
\end{equation}

So, by calculating the death rate $\lambda$, we can find the ground state energy
of a particle in an infinite square well.

But the known ground state of a particle in an infinite square well is
$\frac{\pi^2}{8}\frac{\hbar^2}{mJ^2}$ \todo{cite}. This implies that we expect $\lambda
R^2=\frac{\pi^2}{8}$ where R is the boundary of the well. Thus, we expect to
find a $\lambda$ that is close to this theoretical value.

@@

Thus, by plotting the logarithm of the probability of surviving $n$ steps
against the number of steps $n$, we can obtain the death rate $\lambda$ from the
gradient.

Varying the boundary $J$ of the square well will affect the accuracy of the
final answer, and this is also investigated here.

>>>>

\begin{equation}
  \label{eq:squarewell}
    a(j) =
    \begin{cases}
      0,& \text{if } j < J\\
      \infty,& \text{if } j \geq J
    \end{cases}
    \nonumber
\end{equation}
where J is a well-defined boundary. So the walker cannot die inside the well but
wil immediately die when they step outside, which we expect will definitely
happen given sufficient time. Inside the well, the potential is simply that of a
free particle \todo{cite}:

@@

@@

Thus, by plotting the logarithm of the probability of surviving $n$ steps
against the number of steps $n$, we can obtain the death rate $\lambda$ from the
gradient of the line.

\section{Method}
We implemented the Random Walk simulation computationally in
C++ in order to optimise the performance of the simulation, and thereby allow
for a larger number of iterations and larger potential wells. The source code is
available \cite{Pandian}. The simulation is easily extensible to multiple
dimensions with a linear increase in time-complexity as the number of dimensions
increases, but this investigation focused on a one-dimensional system.

The random direction is selected using a pseudorandom number generator (pRNG)
known as the Mersenne Twister. This algorithm was chosen as it is a very fast
implementation of a pRNG \cite{Route2017}, passes many statistical randomness
tests \cite{Bellamy2013}, and has a very large period of $2^{19937}-1$.

We initially ran the simulation for a boundary of $J = 20$, with $1000000$
walkers. The step on which each walker died was recorded so that a probability
distribution showing the likelihood of dying on a particular step could be
constructed.

We then consider the probability that a walker will survive $n$ steps, and use
the relationship between the number of walkers surviving $n$ steps and $n$ to
determine the rate of death $\lambda$, as shown in Equation \ref{eq:straightline}.

One method of reducing the computing time of the simulation is to introduce a
maximum number of steps. The walker will die if they reach the boundary of the
well, or if they reach this maximum number of steps, whichever happens
first. However, this will introduce some systematic error into the simulation,
as there will be a disproportionately high number of walkers dying on the
maximum number of steps. We also investigated these effects here.

We expect to see an effect on the accuracy of our final ground state depending
on the size of the infinite well. We investigate these effects by comparing the
calculated ground state energy to the theoretical energy for different sized wells.
